---
title: "Clustering Walkthrough"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo = FALSE, eval = TRUE}
library(tidyverse) 
library(cluster) #cluster analysis
library(factoextra) #visualization
library(tidymodels)# just preprocessing 
```


```{r}
# Full ames data set --> recode ordinal variables to numeric
dat <- AmesHousing::make_ames()

ames_num <-  dat %>%
  mutate_if(str_detect(names(.), 'Qual|Cond|QC|Qu'), as.numeric) 
```

First we need to preprocess the data.  Dummy code the nominal variables, normalize all the numeric variables (scale matters here), then prep and bake.
```{r}

ames_recipe <- recipe(Sale_Price ~ .,
                   data = ames_num) |> 
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) |>  # For all the nominal variable
  step_normalize(all_numeric(), -all_outcomes(), ) |> #So they are all in the correct proportion
  prep() %>%  # Apply the steps. Calculates all the steps for nomial variables it in the background
  bake(., ames_num) %>% 
  select(. , -c(Neighborhood_Hayden_Lake, Sale_Price))

```

K-means clustering.  For this algorithm, we need to specify the number of clusters to form.  How many do you think there are?  
```{r}

ames_clust <- kmeans(ames_recipe, centers = 6) #Creates the clusters by spatial caracteristics

summary(ames_clust)
##tidy(ames_clust)

#now let's try a systematic method for setting k

clusters <- fviz_nbclust(ames_recipe,
                         kmeans,
                         method = "wss",
                         k.max = 25, # Number of clusters
                         verbose = FALSE) 
#class(clusters)

# Seek for the nature of the data and look for the "ELBOW" It is important to find that balance between different clusters and SSE.

```

```{r}
#We can examine predictor averages for each cluster

augment(ames_clust, ames_recipe) |> 
  ggplot(aes(Latitude, Longitude, color = .cluster)) +
  geom_point()


augment_check <- augment(ames_clust, ames_recipe)

class(augment_check)

names(augment_check)

class(augment_check$.cluster[1])

```

Update the dataset
```{r}

final_data <- cbind(ames_recipe, cluster = ames_clust$cluster)


```

Hierarchical clustering
Now it's your turn to partition the dataset, this time using hclust().

1. The first thing to do is calculate a distance matrix on the data (using dist()) that contains info on how far apart each observation is from each other observation.  

2.Use tidy() on the distance matrix so you can see what is going on.


3.Then apply the clustering method with hclust().


4.How does the plot look?  For clarity, let's try this again with a subset of our data.  Take a random sample of 100 observations from the data set and run hclust() on that.  Now plot.  Do you see any outliers?  How can you tell?  